 Casting multiple shadows: interactive data visualisation with tours and embeddings

### 2020-02-05 paper.Rmd
more words

## Goals of high dimensional vis

## How do analysts actually use these methods

Broadly characterised by @Brehmer2014-hk, characterised two main workflow
steps that an analyst performs when using embedding methods: dimension reduction
and cluster orientation. The first relates to 
dimension reduction achieved by using an embedding method, here an analyst
wants to characterise and map meaning onto the embedded form, for example
identifying batch effects from a high throughput sequencing experiment, or
identifying a gradient along the embedded form. The second relates to using
embeddings as part of a clustering workflow. Here analysts are interested
in identifying  and naming clusters and verifying them by either applying 
known labels or colouring by variables that are a-priori known to distinguish 
clusters.

@@

>>>>

High dimensional data is increasingly prevalent in
the natural sciences and beyond but presents a challenge to the analyst in terms of both data
cleaning / preprocessing and visualisation. Methods
to embed data from a high-dimesional space into a low-dimensional one now form a core step of the data analysis workflow where they are used to ascertain hidden structure and denoise data for downstream analysis (thereby nullifying the 'curse of dimensionality').

Choosing an appropriate embedding presents a 
challenge to the analyst. How does an analyst know
whether the embedding has captured the underlying
topology and geometry of the high dimensional space?
The answer depends on the analyst's workflow. 
@Brehmer2014-hk characterised two main workflow
steps that an analyst performs when using embedding techniques: dimension reduction
and cluster orientation. The first relates to 
dimension reduction achieved by using an embedding method, here an analyst
wants to characterise and map meaning onto the embedded form, for example
identifying batch effects from a high throughput sequencing experiment, or
identifying a gradient or trajectory along the embedded form. The second relates to using
embeddings as part of a clustering workflow. Here analysts are interested
in identifying  and naming clusters and verifying them by either applying 
known labels or colouring by variables that are a-priori known to distinguish 
clusters. Both of these workflow steps rely on the embedding being 'faithful' or the original high dimensional dataset, and become much more difficult when there is no underlying ground truth.

@@

We discuss the use of

- PCA
- tSNE
- UMAP
- PHATE
